Boosting, Sparsity, and Conditional Gradient: A Tutorial

### 2014-03-27 arXiv/introduction.tex
fixed some small errors
rewrote theorem 2.7 (lower bound)

With these examples in mind, we are now ready to state our main result:
\begin{theorem}
\label{thm:main}
Given vector spaces $\V$ and $\W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $h : \W \to \bR$ be a convex function and let $f : \V \times \W \to \reals$ be a function 
such that
\begin{enumerate}
\item $f(\cdot,w)$ is linear for each fixed $w$
\item $f(v, \cdot) - h(\cdot)$ is convex for each fixed $v$.
\end{enumerate}
Suppose further that $h$ is strongly convex with respect to a norm $\|\cdot\|$. 
\fTBD{JHH: we haven't defined strongly convex yet}
Then, for any compact set $S \subset \V$, and any point $x \in \Conv(S)$, 
there exists an $(S,k)$-sparse point\fTBD{JHH: note that it should be {\bf an} $(S,k)$-sparse point} $y$ such that 
\[ \max_{w \in \W} f(y,w) \leq \max_{w \in \W} f(x,w) + \frac{2D^2}{k+1}, \]
where $D = \sup_{s, s' \in S} \|s-s'\|_*$ and $\|\cdot\|_*$ is the dual norm 
of $\|\cdot\|$.
\end{theorem}
This result is tight in the following sense:\fTBD{JS: can we generalize this to arbitrary norms? or some norms other than $\ell_{2}$?}\fTBD{JHH: rewrite to explicitly give $f$ and $S$}
\begin{theorem}
\label{thm:lower-bound}
There exists a choice of $f$ satisfying the conditions of Theorem~\ref{thm:main}, 
a compact set $S \subset \V$, and a point $x \in \Conv(S)$ such that for any 
$(S,k)$-sparse point $y$, 
\[ \max_{w \in \W} f(y,w) \ge \max_{w \in \W} f(x,w) + \Theta(1/k). \]
\end{theorem}
The proofs of Theorems~\ref{thm:main} and \ref{thm:lower-bound} are given in Section~\ref{TODO}. For now, 
to give a flavor for the relevance of Theorem~\ref{thm:main}, we show how 
to obtain a sparsification result for Example~\ref{ex:marginal-polytope}.
\begin{proposition}
\label{prop:pseudosamples}
Let $\phi : \sX \to \reals^d$ and let $\sM = \Conv(\{\phi(x) \mid x \in \sX\})$ 
be the marginal polytope as defined in Example~\ref{ex:marginal-polytope}. 
Let $D$ be the diameter of $\sM$ in the $l^2$ norm. Then, for any 
$\mu \in \sM$, there exists points $x_1,\ldots,x_k$ and weights 
$w_1,\ldots,w_k$ such that
\[ \left\|\mu - \sum_{i=1}^k w_i \phi(x_i)\right\|_2 \leq \sqrt{\frac{4D^2}{k+1}} \]
\end{proposition}
\begin{proof}
Let $\V = \W = \reals^d$, and apply Theorem~\ref{thm:main} to the function 
$f(v,w) = (v-\mu)^{\top}w - \frac{1}{2}\|w\|_2^2$. The function $f$ is 
linear in $v$ for each $w$, and is strongly convex in $w$ with respect to the 
$l^2$ norm for each $v$. Also note that 
$\sup_{w \in W} f(v,w) = \frac{1}{2}\|v-\mu\|_2^2$. Since the dual of the 
$l^2$ norm is the $l^2$ norm, Theorem~\ref{thm:main} says that there exists 
a $(S,k)$-sparse point $y$ such that
\[ \frac{1}{2}\|\mu - y\|_2^2 \leq \frac{1}{2}\|\mu - \mu\|_2^2 + \frac{2D^2}{k+1}. \]
Since $\|\mu-\mu\|_2 = 0$ and any $(S,k)$-sparse point $y$ can be written 
as $\sum_{i=1}^k w_i \phi(x_i)$, we have the desired result.
\end{proof}
An analogous result can be obtained by instead choosing $\phi : \mcX \to \mcH$, where $(\mcH, \|\cdot\|_{\mcH})$ is a reproducing kernel Hilbert space and using $\|\cdot\|_{\mcH}$ instead of $\|\cdot\|_{2}$. 
Proposition~\ref{prop:pseudosamples} says that, given any distribution 
$\pi$ over $\sX$, we can find $O(1/\epsilon^2)$ pseudosamples that approximate 
the moments of $\pi$ with accuracy $\epsilon$ in the $l^2$ (or Hilbert) norm.\fTBD{JHH: added Hilbert space discussion} As
we shall see in Section \ref{sec:algorithm}, in some instances this 
can be improved to $O(1/\epsilon)$.\fTBD{JHH: will give Chen extension, so no need
to defer to Chen paper here}
%We note that in some instances this can be improved to $O(1/\epsilon)$, see 
%for instance \cite{chen}.

@@

\NA{Proposed sections:
\begin{enumerate}
\item (B) Intro
\item (B) Sparsity result and lower bound
\item (JS) BMD and proofs of sparsity result and lower bound
\item (JH) Application: Low-rank matrix approximation, approximation of SDP matrix sums 
\item (JH) Application: Graph sparsification, max flow
\item Application: Solving games 
\item (JH) Application: Herding, including Chen extension
\item (JS) Application: Circuits and boosting
\item Concluding thoughts 
\end{enumerate}
\qquad(A) Appendix: BMD proofs}

>>>>

With these examples in mind, we are now ready to state our main result:
\begin{theorem}
\label{thm:main}
Given vector spaces $\V$ and $\W$\fTBD{JHH: hmm...you've switched from $\reals^{d}$ to vector spaces}, let $h : \W \to \reals$ be a convex function and let $f : \V \times \W \to \reals$ be a function 
such that
\begin{enumerate}
\item $f(\cdot,w)$ is linear for each fixed $w$
\item $f(v, \cdot) - h(\cdot)$ is convex for each fixed $v$.
\end{enumerate}
Suppose further that $h$ is strongly convex with respect to a norm $\|\cdot\|$. 
\fTBD{JHH: we haven't defined strongly convex yet}
Then, for any compact set $S \subset \V$, and any point $x \in \Conv(S)$, 
there exists an $(S,k)$-sparse point\fTBD{JHH: note that it should be {\bf an} $(S,k)$-sparse point} $y$ such that 
\[ \max_{w \in \W} f(y,w) \leq \max_{w \in \W} f(x,w) + \frac{2D^2}{k+1}, \]
where $D = \sup_{s, s' \in S} \|s-s'\|_*$ and $\|\cdot\|_*$ is the dual norm 
of $\|\cdot\|$.
\end{theorem}
This result is tight in the following sense:\fTBD{JS: can we generalize this to arbitrary norms? or some norms other than $\ell_{2}$?}
\begin{theorem}
\label{thm:lower-bound}
Choose $S = \Delta_{n}$ as in Example \ref{ex:simplex} with $\V = \W = \Conv(S)$. Let $f(v,w) = \ip{v}{w} - \frac{1}{2}\|w\|^{2}$ (which satisfies the conditions of Theorem~\ref{thm:main}). Then with $x = 0$, for any 
$(S,k)$-sparse point $y$, 
\[ \max_{w \in \W} f(y,w) \ge \max_{w \in \W} f(x,w) + \Theta(1/k), \]
i.e.,
\[ \frac{1}{2}\|y\|_{2}^{2} \ge \Theta(1/k). \]
\end{theorem}
The proofs of Theorems~\ref{thm:main} and \ref{thm:lower-bound} are given in Section~\ref{TODO}. For now, 
to give a flavor for the relevance of Theorem~\ref{thm:main}, we show how 
to obtain a sparsification result for Example~\ref{ex:marginal-polytope}.
\begin{proposition}
\label{prop:pseudosamples}
Let $\phi : \sX \to \reals^d$ and let $\sM = \Conv(\{\phi(x) \mid x \in \sX\})$ 
be the marginal polytope as defined in Example~\ref{ex:marginal-polytope}. 
Let $D$ be the diameter of $\sM$ in the $l^2$ norm. Then, for any 
$\mu \in \sM$, there exists points $x_1,\ldots,x_k$ and weights 
$w_1,\ldots,w_k$ such that
\[ \left\|\mu - \sum_{i=1}^k w_i \phi(x_i)\right\|_2 \leq \sqrt{\frac{4D^2}{k+1}} \]
\end{proposition}
\begin{proof}
Let $\V = \W = \reals^d$, and apply Theorem~\ref{thm:main} to the function 
$f(v,w) = (v-\mu)^{\top}w - \frac{1}{2}\|w\|_2^2$. The function $f$ is 
linear in $v$ for each $w$, and is strongly convex in $w$ with respect to the 
$l^2$ norm for each $v$. Also note that 
$\sup_{w \in W} f(v,w) = \frac{1}{2}\|v-\mu\|_2^2$. Since the dual of the 
$l^2$ norm is the $l^2$ norm, Theorem~\ref{thm:main} says that there exists 
a $(S,k)$-sparse point $y$ such that
\[ \frac{1}{2}\|\mu - y\|_2^2 \leq \frac{1}{2}\|\mu - \mu\|_2^2 + \frac{2D^2}{k+1}. \]
Since $\|\mu-\mu\|_2 = 0$ and any $(S,k)$-sparse point $y$ can be written 
as $\sum_{i=1}^k w_i \phi(x_i)$, we have the desired result.
\end{proof}
An analogous result can be obtained by instead choosing $\phi : \mcX \to \mcH$, where $(\mcH, \|\cdot\|_{\mcH})$ is a reproducing kernel Hilbert space and using $\|\cdot\|_{\mcH}$ instead of $\|\cdot\|_{2}$. 
Proposition~\ref{prop:pseudosamples} says that, given any distribution 
$\pi$ over $\sX$, we can find $O(1/\epsilon^2)$ pseudosamples that approximate 
the moments of $\pi$ with accuracy $\epsilon$ in the $l^2$ (or Hilbert) norm.\fTBD{JHH: added Hilbert space discussion} As
we shall see in Section \ref{sec:algorithm}, in some instances this 
can be improved to $O(1/\epsilon)$.\fTBD{JHH: will give Chen extension, so no need
to defer to Chen paper here}
%We note that in some instances this can be improved to $O(1/\epsilon)$, see 
%for instance \cite{chen}.

@@

\NA{Proposed sections:
\begin{enumerate}
\item (B) Intro
\item (B) Sparsity result and lower bound
\item (JS) BMD and proofs of sparsity result and lower bound
\item (JH) Application: Low-rank matrix approximation, approximation of PSD matrix sums 
\item (JH) Application: Graph sparsification, max flow
\item Application: Solving games 
\item (JH) Application: Herding, including Chen extension
\item (JS) Application: Circuits and boosting
\item Concluding thoughts 
\end{enumerate}
\qquad(A) Appendix: BMD proofs}
